---
title: "Assignment 3"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Text and Network Visualization with ArXiv Papers (Part 2)

## Author Network Analysis (2)
```{r}
# setup 
rm(list = ls())

# libraries
library(tidyverse) 
library(igraph)
library(visNetwork)

# loading data
arxiv <- read_csv("data/arxiv_subset.csv")
```

### a) **Wrangling & Network Construction**

- **Filter the dataset** to include only papers that are categorized as both "physics.soc-ph" (Physics and Society) AND "cs.SI" (Social and Information Networks)
- Extract author information from these papers and create a network where authors who co-authored at least one paper are connected
  - _Hint: For each paper, create pairs of co-authors. These pairs will form the edges of your network._
- **Further limit your analysis to authors who have published at least 4 papers** in this subset to keep the visualization manageable
  - _Hint: Count papers per author, then filter to keep only prolific authors._

```{r}
#wrangling 
filtered <- arxiv %>%
  filter(str_detect(categories, "physics.soc-ph") & str_detect(categories, "cs.SI")) %>%
  separate_rows(authors, sep = ",") %>%
  separate_rows(authors, sep = " and ") %>%
  mutate(authors = str_trim(authors)) %>%
  filter(authors != "")

# Recognizing accent marks
filtered <- filtered %>%
  mutate(
    # Acute accents (á, é, etc.)
    authors = str_replace_all(authors, fixed("\\'a"), "á"),
    authors = str_replace_all(authors, fixed("\\'e"), "é"),
    authors = str_replace_all(authors, fixed("\\'i"), "í"),
    authors = str_replace_all(authors, fixed("\\'o"), "ó"),
    authors = str_replace_all(authors, fixed("\\'u"), "ú"),
    
    # Grave accents (à, è, etc.)
    authors = str_replace_all(authors, fixed("\\`a"), "à"),
    authors = str_replace_all(authors, fixed("\\`e"), "è"),
    authors = str_replace_all(authors, fixed("\\`i"), "ì"),
    authors = str_replace_all(authors, fixed("\\`o"), "ò"),
    authors = str_replace_all(authors, fixed("\\`u"), "ù"),
    
    # Umlaut/diaeresis (ä, ö, etc.)
    authors = str_replace_all(authors, fixed('\\"a'), "ä"),
    authors = str_replace_all(authors, fixed('\\"o'), "ö"),
    authors = str_replace_all(authors, fixed('\\"u'), "ü"),
    
    # Circumflex (â, ê, etc.)
    authors = str_replace_all(authors, fixed("\\^a"), "â"),
    authors = str_replace_all(authors, fixed("\\^e"), "ê"),
    authors = str_replace_all(authors, fixed("\\^i"), "î"),
    authors = str_replace_all(authors, fixed("\\^o"), "ô"),
    authors = str_replace_all(authors, fixed("\\^u"), "û"),
    
    # Tilde (ñ, Ñ)
    authors = str_replace_all(authors, fixed("\\~n"), "ñ"),
    authors = str_replace_all(authors, fixed("\\~N"), "Ñ"),
    
    # Cedilla (ç, Ç)
    authors = str_replace_all(authors, fixed("\\c{c}"), "ç"),
    authors = str_replace_all(authors, fixed("\\c{C}"), "Ç"),
    
    # Remove any remaining backslashes
    authors = str_replace_all(authors, fixed("\\"), "")
  )

#paper_counts DF 
paper_counts_df <- filtered %>%
  count(authors, name = "paper_count")

#converting var to numeric in paper_counts DF 
paper_counts_df$paper_count <- as.numeric(paper_counts_df$paper_count)

#merging paper_counts to filtered
filtered <- filtered %>% 
  left_join(paper_counts_df, by = 'authors')

# further filtering to authors w/ more than 4 papers
prolific_authors <- filtered %>%
  filter(paper_count >= 4)

# creating vector 
prolific_authors_vect <- prolific_authors %>%
  pull(authors)

#edges 
edges <- filtered %>%
  group_by(id) %>%
  mutate(prolific_in_paper = authors %in% prolific_authors_vect) %>%
  filter(prolific_in_paper) %>%
  filter(n() >= 2) %>%
  summarise(authors = list(unique(authors)), .groups = "drop") %>%
  mutate(pairs = purrr::map(authors, ~ combn(.x, 2, simplify = FALSE))) %>%
  unnest(pairs) %>%
  mutate(
    from = map_chr(pairs, ~ .x[1]),
    to = map_chr(pairs, ~ .x[2])
  ) %>%
  select(from, to)

#adding weights based on paper number 
edges_weighted <- edges %>%
  count(from, to, name = "weight")

#output 
g <- graph_from_data_frame(edges_weighted, directed = FALSE)

#checking NAs for papers 
V(g)$papers <- paper_counts_df$paper_count[match(V(g)$name, paper_counts_df$authors)]
if (any(is.na(V(g)$papers))) {
  V(g)$papers[is.na(V(g)$papers)] <- 0
}

```

### b) **Preparing Network for Visualization**

- Create a visualization of the author collaboration network using an appropriate layout algorithm
- Size nodes based on the number of papers published by each author
  - _Hint: Map the paper count to node size in your visualization._
- Color nodes based on a relevant metric (e.g., betweenness centrality or degree)
  - _Hint: Use centrality metrics available in network analysis packages like `igraph`._
- Label the 10-15 most prolific authors

```{r}
# Centrality Calculation 
V(g)$betweenness <- betweenness(g, normalized = TRUE)

# Top Authors for Labeling 
top_authors <- V(g)$name[order(V(g)$papers, decreasing = TRUE)][1:15]

# Nodes and Edges used for visNetwork package
nodes_df <- data.frame(
  id = V(g)$name,
  label = ifelse(V(g)$name %in% top_authors, V(g)$name, ""),
  value = V(g)$papers,
  color = colorRampPalette(c("#54C568FF", "#36A9ABFF", "#3487A6FF", "#3D5296FF"))(100)[
    cut(V(g)$betweenness, breaks = 400, labels = FALSE)
  ],
  title = paste0(
    "<b>", V(g)$name, "</b><br>",
    "Papers: ", V(g)$papers, "<br>",
    "Betweenness: ", round(V(g)$betweenness, 3)
  )
)

#df for edges
edges_df <- edges_weighted %>%
  select(from, to)
```

## Visualizing Network
```{r}
# Interactive Network Visualization 
author_net <- visNetwork(nodes_df, edges_df, height = "600px", width = "100%") %>%
  visIgraphLayout(layout = "layout_nicely") %>%
  visNodes(size = nodes_df$value, font = list(size = 45, color = "#444444", face = "georgia", bold = TRUE)) %>%
  visEdges(smooth = FALSE) %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>% 
  visInteraction(keyboard = TRUE, dragNodes = TRUE, dragView = TRUE, zoomView = TRUE) %>%
  visLayout(randomSeed = 42) %>% 
  visEvents(
    type = "once",
    startStabilizing = "function() { this.moveTo({scale:0.15}); }")

author_net

# save 
visSave(author_net, file = "figures/author_network.html")

```
In the network displayed, nodes represent authors and edges (connections) indicate co-authorship relationships. The size of nodes represent the number of papers published by each author, while node color reflects centrality measure of betweenness, with with darker blue highlighting more central or influential authors. The network exhibits a densely connected core with several prominent and well-connected authors labeled (top 25). They are surrounded by peripheral nodes and smaller, loosely connected clusters as we go further from the center. 


