---
title: "Assignment 3.1 Text Analysis"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Text and Network Visualization with ArXiv Papers (Part 1)

**Data**
ArXiv is a free distribution service and open-access archive for scholarly articles, primarily in physics, mathematics, computer science, and related disciplines. For this assignment, we'll be working with a dataset of recent papers from the "Physics and Society" (physics.soc-ph) category, which covers research on social phenomena, collective behavior, opinion dynamics, social networks, and other topics at the intersection of physics and social science.

The dataset contains information about approximately 10,000 recent papers published between 2018 and the present. This subset has been extracted from the complete ArXiv dataset.

**Variable descriptions:** `arxiv_subset.csv`

| variable    | class     | description                                                 |
| :---------- | :-------- | :---------------------------------------------------------- |
| id          | character | ArXiv paper ID                                              |
| title       | character | Paper title                                                 |
| authors     | character | Author names (comma-separated list)                         |
| categories  | character | ArXiv categories (papers can belong to multiple categories) |
| update_date | character | Last update date                                            |
| abstract    | character | Full text abstract of the paper                             |

Additional information is available in the full JSON file (`arxiv_subset.json`) including version history, DOI, and other metadata. The JSON file also contains the parsed author names, making it easier to work with author data.

**Tasks**
In this assignment, you will create a series of visualizations to explore the ArXiv papers dataset through text and network analysis techniques. Your notebook should be well-documented with explanations of your approach and interpretation of results. *Choose two of the three main tasks below to complete* (***Text Analysis, Author Network***, or Topic Analysis).

## Term Frequency Analysis (1)
```{r}
#setup 
rm(list = ls())

# libraries
library(tidyverse)
library(tidytext)
library(lubridate)
library(wordcloud)
library(textstem)
#library(SnowballC)

# loading data
arxiv <- read_csv("data/arxiv_subset.csv")
```

## Part 1a: Term Frequency, setup 
- Prepare the paper abstracts by removing stopwords, punctuation, and numbers
- _Hint: In R, the `tidytext` package provides functions like `unnest_tokens()` to tokenize text; in Python, use `nltk` or `spaCy` for tokenization and stopword removal._
- Create a bar chart showing the top 20 most frequent terms across all abstracts
- _Hint: After tokenizing, count term frequencies and sort to find the most common terms._
- Generate a comparative visualization (e.g., small multiples or grouped bars) showing how the top 10 terms differ between the years 2019 and 2023
- _Hint: Use `update_date` as the source of the paper year._

```{r}
# Tokenizing and cleaning abstracts
abstract_words <- arxiv %>%
  select(id, abstract, update_date) %>%
  unnest_tokens(word, abstract) %>% # Splits into words
  filter(!str_detect(word, "\\d")) %>% # Removes numbers
  anti_join(get_stopwords()) %>% # Removes stopwords
  filter(!word %in% c("’", "“", "”", "–")) %>%  # Removes punctuation attached to words
  mutate(word = lemmatize_words(word)) #lemmatizes words (matches words to root word)

# Identifying top 20 terms across all abstracts (2018-2023)
top_terms_all <- abstract_words %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20)

# Plot:  top 20 terms across all abstracts (all years)
ggplot(top_terms_all, aes(x = reorder(word, n), y = n, fill = n)) + #descending order 
  geom_col() +
  scale_fill_gradient(low = "#6878ad", high = "#3D5296FF") +
  coord_flip() +
  labs(x = NULL, y = 'Counts', 
       title = 'Top 20 Most Frequent Words', 
       subtitle = 'Analyzing Abstracts from ArXiv Journal Publications (2018-2023)') +
  theme_minimal() + theme(legend.position="none")

# save
ggsave("figures/term_freq_all_years.png", width = 8, height = 6)
```

Note: I initially applied the Snowball package in an attempt to handle words that contain the same root but different ending (e.g., network and networks). The output of stemming, which is the process of truncating words to roots, resulted in an output of non-real words (propos, measur, etc.). I then tried lemmatization using the textstem package, designed to convert words to their dictionary stem-word (lemma) using linguistic rules (which I will not talk about). Inputting this into my pipeline (seen above) resolved the issue because there are no longer multiple variations of the same word or stems of words (which aren't words). 


```{r}
# Filters to words from 2019 and 2023 only
abstract_words_19_23 <- abstract_words %>%
  mutate(year = year(update_date)) %>% # extracts only year in 'update_date' column
  filter(year %in% c(2019, 2023)) # filters for only 2019 and 2023

#DF term frequency: counts words from these years
term_freq_year <- abstract_words_19_23 %>%
  count(year, word, sort = TRUE)

# Vector: gets top 10 terms specifically aggregated from 2019 and 2023
top_terms_19_23 <- term_freq_year %>%
  group_by(word) %>%
  summarise(total = sum(n)) %>%
  slice_max(total, n = 10) %>% #top 10 words 
  pull(word)

# Filter to these top terms
comparison_data <- term_freq_year %>%
  filter(word %in% top_terms_19_23) %>%
  complete(year, word, fill = list(n = 0))   # Ensure all terms appear in both years

#Plot: top 10 words 
ggplot(comparison_data, aes(x = reorder(word, n), y = n, fill = factor(year))) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("2019" = "#3D5296FF", "2023" = "#36A9ABFF")) +
  coord_flip() +
  labs(x = NULL, y = 'Counts', 
       title = 'Top 10 Most Frequent Words (2019 vs. 2023)', 
       subtitle = 'Analyzing Abstracts from ArXiv Journal Publications') +
  theme_minimal()

#save 
ggsave("figures/term_freq_2019_2023.png", width = 8, height = 6)
```
This analysis, clearly shows a significant shift in the popularity of words used in abstracts between the years 2019 and 2023. When we aggregate and count the frequency of words across all abstracts, regardless of year, and then identify the top 10 words, we notice great variation. Based on word frequency across allyears, we found that 2023 contained more of the most frequently used words across abstracts compared to 2019. This suggests that 2023 articles contain more words that are more commonly used in journal articles within the years 2018-2023 compared to those of 2019. Visually, it seems there is around a 0.5 (0.4-0.6) difference in word use across the top 10 words. 

b) **Word Cloud**
- Option 1: Generate a word cloud of the 100 most frequent meaningful terms in the abstracts, with proper color encoding and layout.
- _Hint: Use the `wordcloud` package in R to create visualizations with control over colors and sizes._

## 1b: Word Cloud
```{r}
# for reproducibility
set.seed(128)

#setting up
top_100 <- abstract_words %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 100)

#plotting & saving
png("figures/wordcloud_top_100.png", width = 600, height = 400)
wordcloud(top_100$word, top_100$n,
          scale = c(5, 0.5), colors = viridis::viridis(100))

dev.off()
```
Based on the word cloud, we also notice that the words 'network' and 'model' are highly distinquishable compared to all other top 100 words. Interestingly, we this reflected in both bar graphs, showing a large difference in the rate of frequency between the words 'network' and 'model' and all other words. 
